\documentclass[paper]{JHEP3}
\usepackage{amsmath,amssymb,enumerate,url}
\bibliographystyle{JHEP}

%%%%%%%%%% Start TeXmacs macros
\newcommand{\tmtextit}[1]{{\itshape{#1}}}
\newcommand{\tmtexttt}[1]{{\ttfamily{#1}}}
\newenvironment{enumeratenumeric}{\begin{enumerate}[1.] }{\end{enumerate}}
\newcommand\sss{\mathchoice%
{\displaystyle}%
{\scriptstyle}%
{\scriptscriptstyle}%
{\scriptscriptstyle}%
}

\newcommand\as{\alpha_{\sss\rm S}}
\newcommand\POWHEG{{\tt POWHEG}}
\newcommand\POWHEGBOX{{\tt POWHEG BOX}}
\newcommand\HERWIG{{\tt HERWIG}}
\newcommand\PYTHIA{{\tt PYTHIA}}
\newcommand\MCatNLO{{\tt MC@NLO}}

\newcommand\kt{k_{\sss\rm T}}

\newcommand\pt{p_{\sss\rm T}}
\newcommand\LambdaQCD{\Lambda_{\scriptscriptstyle QCD}}
%%%%%%%%%% End TeXmacs macros


\title{The POWHEG BOX user manual:\\
  Higgs boson production through  gluon fusion in the 2HDM} \vfill
\author{Simone Alioli\\
  Deutsches Elektronen-Synchrotron DESY\\
  Platanenallee 6, D-15738 Zeuthen, Germany\\
  E-mail: \email{simone.alioli@desy.de}}
\author{Paolo Nason\\
  INFN, Sezione di Milano-Bicocca,
  Piazza della Scienza 3, 20126 Milan, Italy\\
  E-mail: \email{Paolo.Nason@mib.infn.it}}
\author{Carlo Oleari\\
  Universit\`a di Milano-Bicocca and INFN, Sezione di Milano-Bicocca\\
  Piazza della Scienza 3, 20126 Milan, Italy\\
  E-mail: \email{Carlo.Oleari@mib.infn.it}}
\author{Emanuele Re\\
  Institute for Particle Physics Phenomenology, Department of Physics\\
  University of Durham, Durham, DH1 3LE, UK\\
  E-mail: \email{emanuele.re@durham.ac.uk}}
\author{Emanuele Angelo Bagnaschi, Pietro Slavich\\
  LPTHE, 4, Place Jussieu, F-75252 Paris, France\\
  E-mail: \email{bagnaschi@lpthe.jussieu.fr, slavich@lpthe.jussieu.fr}}
\author{Giuseppe Degrassi\\
  Dipartimento di Fisica, Universit\`a di Roma Tre and INFN, sezione di Roma Tre\\
  Via della Vasca Navale 84, I-00146, Rome, Italy\\
  E-mail: \email{degrassi@fis.uniroma3.it}}
\author{Alessandro Vicini\\
  Dipartimento di Fisica, Universit\`a di Milano and INFN, sezione di Milano\\
  Via Celoria 16, I-20133 Milano, Italy\\
  E-mail: \email{alessandro.vicini@mi.infn.it}}

\vskip -0.5truecm

\keywords{POWHEG, Shower Monte Carlo, NLO}

\abstract{This note documents the use of the package
  \tmtexttt{POWHEG BOX} for Higgs boson production trough gluon fusion
  in the 2HDM.
  Results can be easily interfaced to shower Monte Carlo programs, in
  such a way that both NLO and shower accuracy are maintained.}
\preprint{\today\\ \tmtexttt{POWHEG BOX} 1.99-beta}

\begin{document}


\section{Introduction}

The \tmtexttt{POWHEG BOX} program is a framework for implementing NLO
calculations in Shower Monte Carlo programs according to the
\POWHEG{} method. An explanation of the method and a discussion of
how the code is organized can be found in
refs.~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}.  The code is
distributed according to the ``MCNET GUIDELINES for Event Generator
Authors and Users'' and can be found at the web page \\
%
\begin{center}
 \url{http://powhegbox.mib.infn.it}.
\end{center}
%
~\\
%


This program is an implementation of the NLO cross section for Higgs boson production via gluon fusion process, in the MSSM,
in the POWHEG formalism of refs. \cite{Nason:2004rx,Frixione:2007vw}. 
It is based over the work done by \emph{Alioli et al} ~\cite{Alioli:2008tz}.
The supported processes are the production of the light CP even scalar
$h$, of the heavy CP even scalar $H$ and of the pseudoscalar $A$.
This implementation provides the exact treatment  in the amplitude of the quark
and of the squark masses; it also provides the evaluation of
the diagrams involving the quark-squark-gluino vertices, which are
approximated by different mass expansions, as described in refs.\cite{Degrassi:2008zj, Degrassi:2010eu, Degrassi:2011vq}.
MSSM parameters are 
A detailed description of the implementation can be found on ref.~\cite{Bagnaschi:2011tu}.


Spin correlations of Higgs boson decay products are not included, being it
a scalar. This issue can be safely left to the subsequent Shower Monte
Carlo program. Finite Higgs boson width effects are accounted for.

The code, that can be found in the \tmtexttt{POWHEG-BOX/gg\_H\_MSSM}
subdirectory is based on the subtraction scheme by Frixione, Kunszt
and Signer implemented in the \tmtexttt{POWHEG BOX}, rather than on the scheme
discussed in the paper~\cite{Alioli:2008tz}. Please cite it anyhow if
you use the program.

In order to run the \tmtexttt{POWHEG BOX} program, we recommend the
reader to start from the \tmtexttt{POWHEG BOX} user manual, which
contains all the information and settings that are common between
all subprocesses. In this note we focus on  the settings and
parameters specific to $gg\to \phi~~ (\phi=h, H, A)$ implementation.

\section{System requirements}

Chaplin >= 1.2 is required to compile and use this MC event generator.

\section{Generation of events}

Build the executable\\
\tmtexttt{\$ cd POWHEG-BOX/gg\_H\_2HDM \\
\$ make pwhg\_main }\\
Then do (for example) \\
\tmtexttt{\$
cd testrun-lhc\\
\$ ../pwhg\_main}\\
At
the end of the run, the file \tmtexttt{pwgevents.lhe} will contain
100000 events for $gg\to H$  in the Les Houches format.\\~\\
In order to shower them with \PYTHIA{} (for example) do\\
\noindent
\tmtexttt{\$
cd POWHEG-BOX/gg\_H\_2HDM \\ \$ make
main-PYTHIA-lhef \\ \$ cd
testrun-lhc \\
\$ ../main-PYTHIA-lhef}

\section{Process specific input parameters}

The input parameters for POWHEG are separated into two files: one is the standard powheg.input and the others differs according
to the renormalization scheme selected for the MSSM parameters. Currently we implement an On-Shell and a $\overline{\text{DR}}$ scheme.

\subsection{Common configuration file: powheg.input}

The  common parameters, defined in 'powheg.input' are: \\
{\tt
zerowidth 1 ! Control if the Higgs boson is to be produced on-shell or not: 1 = On-Shell; 0 = Off-shell with Breit-Wigner\\
hwidth 10   ! Higgs width, dummy value, you should change it if you do off-shell production\\
bwhape 1       ! BW shape: 1 = running width; 2 = fixed width\\
masswindow 10d0  !(default 10d0) number of widths around hmass in the BW for an off-shell Higgs boson\\
ew 0                     ! ew = 0 disable EW corrections, ew = 1 enable EW corrections\\
fastew 1                   ! fast ew corrections evaluations by mass sampling\\
gfermi 0.116637D-04        ! GF\\
hdecaywidth 0   ! use HDECAY computed width\\
hdecaymode -1      ! PDG code for first decay product of the higgs\\
! allowed values are:  0 all decay channels open\\
!                      1-6 d dbar, u ubar,..., t tbar (as in HERWIG)\\
!                      7-9 e+ e-, mu+ mu-, tau+ tau-\\
!                      10  W+W-\\
!                      11  ZZ\\
!                      12  gamma gamma\\
runningscale 0 ! Use running scale\\
! **** Mandatory parameters for the 2HDM model ****\\
higgstype 1 ! CP-even, neutral, higgs type: 1 = light higgs; 2 = heavy higgs\\
scheme 0    ! Renormalization scheme. 0 = OS , 1 = MSBAR; 2 = DRBAR
topmass 172.5        ! top quark mass
bottommass 4.75d0    ! bottom quark mass - if defined it enables the bottom quark
!charmmass 1.5d0     ! char quark mass - if defined it enables the charm quark
! Optional
! 2HDM parameter
2HDMtype 1           ! 2HDM type; 1=type I; 2=type II; 3=Lepton specific; 4 =Flipped
! decoupling
tanb 10d0            ! Tan beta
alpha -0.09887d0     ! Higgs mixing angle
}

\subsection{General considerations}

The running of $\as$ is evaluated at two loop order, correctly matching,
at flavour thresholds, different definitions that depends on the number of
flavours that can be considered light at the renormalization scale.\\
An example of \tmtexttt{powheg.input} files is given in the subdirectory
 \tmtexttt{gg\_H\_2HDM/testrun-lhc}.
In all examples, the choice of the parameters that control the grid
generation is such that a reasonably small fraction of negative weights is generated, so they
can be run as they are. 
We remind the reader that these negative weights are
only due to our choice of generating $\tilde{B}$ instead of $\bar B$.
They indeed correspond to phase
space points where NLO corrections are bigger than LO contributions. 
Had we performed the integration over the full radiation phase space these negative weights would have
disappeared completely. 

In case one is interfacing to \HERWIG{} or \PYTHIA{} SMC programs, we provide a facility to select the Higgs boson decay products in  these programs :  
~\\
\tmtexttt{
  hdecaymode 12 \phantom{aa} ! code for selection of Higgs boson decay products:\\
\phantom{aaaaaaaaaaaaaaaaa}!  -1 the Higgs boson is left undecayed by the SMC\\ 
\phantom{aaaaaaaaaaaaaaaaa}!  0 all decay channels are open\\
\phantom{aaaaaaaaaaaaaaaaa}!  1-6 d dbar, u ubar,..., t tbar (as in HERWIG)\\
\phantom{aaaaaaaaaaaaaaaaa}!  7-9 e+ e-, mu+ mu-, tau+ tau-\\
\phantom{aaaaaaaaaaaaaaaaa}!  10, 11, 12  W+W-, ZZ, gamma gamma \\     
}
\\

Together with the mandatory parameters, the \POWHEGBOX{} input facility  allows for an easy setting of run parameters, by explicitly adding the relevant lines to the input card. In case one of the following entries is not present in the input card  the reported default value is assumed. In any case, these parameters are printed in the output of the program, so their values can be easily tracked down. 
\\
\tmtexttt{
masswindow 10d0 !(default 10d0) number of widths around hmass in the BW for an off-shell Higgs boson \\
runningscale 0  ! choice for ren and fac scales in Bbar integration}\\
\phantom{pppppppppppppppppppppppppppppp} \tmtexttt{ 0: fixed scale M\_H}\\
\phantom{pppppppppppppppppppppppppppppp} \tmtexttt{ 1: running scale inv mass H}\\ 

Of particular importance are the following parameters: 
\begin{itemize}

\item{\tmtexttt{hhfact    100d0    ! (default no dumping factor) dump factor for high-pt radiation: > 0 dumpfac=h**2/(pt2+h**2) } controls how much of real contribution 
    enters in the \POWHEG{} Sudakov form factor. By default all real
    contributions are included, but this may lead to a NNLO mismatch in the
    higher Higgs boson $\pt$ distribution tail, with respect to fixed order NLO
    results. This actually brings \POWHEGBOX{} results closer to NNLO ones, but if
    one want to switch-off this feature it's possible to use a reduced real
    contribution $R^{\scriptscriptstyle{\texttt{red}}}= R\times$ \tmtexttt{dumpfact} in the Sudakov and to generate the
    remaining $R\times$(1-\tmtexttt{dumpfact}) part without suppression, as documented in Sec. 4.3 of
    ref.~\cite{Alioli:2008tz}.
   }

\item{\tmtexttt{zerowidth 1} (default 0 = false) enforce the calculation in the Higgs zero width approximation.
}

\item{\tmtexttt{bwshape 1} (default 1) choose the functional form of the Breit-Wigner along which the Higgs virtuality is distributed, in case of the zero width approximation  has not been chosen. Allowed values are $1$ for a BW with a running width, $2$ for a fixed width.
}

\end{itemize}

\begin{thebibliography}{10}

\bibitem{Nason:2004rx}
  P.~Nason,
``A new method for combining NLO QCD with shower Monte Carlo algorithms,''
  JHEP {\bf 0411} (2004) 040
  [arXiv:hep-ph/0409146].
  %%CITATION = JHEPA,0411,040;%%

%\cite{Frixione:2007vw}
\bibitem{Frixione:2007vw}
  S.~Frixione, P.~Nason and C.~Oleari,
``Matching NLO QCD computations with Parton Shower simulations: the POWHEG
method,''
  JHEP {\bf 0711} (2007) 070
  [arXiv:0709.2092 [hep-ph]].
  %%CITATION = JHEPA,0711,070;%%


%\cite{Alioli:2010xd}
\bibitem{Alioli:2010xd}
  S.~Alioli, P.~Nason, C.~Oleari and E.~Re,
``A general framework for implementing NLO calculations in shower Monte Carlo
  programs: the POWHEG BOX,''
  [arXiv:1002.2581 [hep-ph]].
  %%CITATION = ARXIV:1002.2581;%%

%\cite{Frixione:2002ik}
\bibitem{Frixione:2002ik}
  S.~Frixione and B.~R.~Webber,
 ``Matching NLO QCD computations and parton shower simulations,''
  JHEP {\bf 0206} (2002) 029
  [arXiv:hep-ph/0204244].
  %%CITATION = JHEPA,0206,029;%%

%\cite{Frixione:2006gn}
\bibitem{Frixione:2006gn}
  S.~Frixione and B.~R.~Webber,
  ``The MC@NLO 3.3 event generator,''
  arXiv:hep-ph/0612272.
%  %%CITATION = HEP-PH/0612272;%%

\bibitem{Dawson:1990zj}
S.~Dawson, {\it {Radiative corrections to Higgs boson production}},  {\em Nucl.
  Phys.} {\bf B359} (1991) 283--300.

\bibitem{Djouadi:1991tka}
A.~Djouadi, M.~Spira, and P.~M. Zerwas, {\it {Production of Higgs bosons in
  proton colliders: QCD corrections}},  {\em Phys. Lett.} {\bf B264} (1991)
  440--446.

\bibitem{Spira:1995rr}
M.~Spira, A.~Djouadi, D.~Graudenz, and P.~M. Zerwas, {\it {Higgs boson
  production at the LHC}},  {\em Nucl. Phys.} {\bf B453} (1995) 17--82,
  [\href{http://xxx.lanl.gov/abs/hep-ph/9504378}{{\tt hep-ph/9504378}}].

%\cite{Alioli:2008tz}
\bibitem{Alioli:2008tz}
  S.~Alioli, P.~Nason, C.~Oleari and E.~Re,
``NLO Higgs boson production via gluon fusion matched with shower in
  POWHEG,''
  arXiv:0812.0578 [hep-ph].
  %%CITATION = ARXIV:0812.0578;%%

\bibitem{mcfm}
  % citazione nostro lavoro
  http://mcfm.fnal.gov/


\bibitem{Cacciari:2005hq}
M.~Cacciari and G.~P. Salam, {\it {Dispelling the $N^3$ myth for the $k_T$
  jet-finder}},  {\em Phys. Lett.} {\bf B641} (2006) 57--61,
  [\href{http://xxx.lanl.gov/abs/hep-ph/0512210}{{\tt hep-ph/0512210}}].


%\cite{Boos:2001cv}
\bibitem{Boos:2001cv}
  E.~Boos {\it et al.},
  ``Generic user process interface for event generators,''
  arXiv:hep-ph/0109068.
  %%CITATION = HEP-PH/0109068;%%

%\cite{Alwall:2006yp}
\bibitem{Alwall:2006yp}
  J.~Alwall {\it et al.},
  ``A standard format for Les Houches event files,''
  Comput.\ Phys.\ Commun.\  {\bf 176} (2007) 300
  [arXiv:hep-ph/0609017].
  %%CITATION = CPHCB,176,300;%%


%\cite{Altarelli:1989wu}
\bibitem{Altarelli:1989wu} T. Sj\"ostrand et~al., in
  ``Z physics at LEP1: Event generators and software,'',  eds.
  G.~Altarelli, R.~Kleiss and C.~Verzegnassi, Vol 3, pg. 327.
  %%CITATION = CERN-89-08-V-3;%%


%\cite{Whalley:2005nh}
\bibitem{Whalley:2005nh}
  M.~R.~Whalley, D.~Bourilkov and R.~C.~Group,
  ``The Les Houches accord PDFs (LHAPDF) and LHAGLUE,''
  arXiv:hep-ph/0508110.
  %%CITATION = HEP-PH/0508110;%%


\bibitem{Alioli:2008gx}
  S.~Alioli, P.~Nason, C.~Oleari and E.~Re,
  ``NLO vector-boson production matched with shower in POWHEG,''
  JHEP {\bf 0807}, 060 (2008)
  [arXiv:0805.4802 [hep-ph]].
  %%CITATION = JHEPA,0807,060;%%

%\cite{arXiv:1111.2854}
\bibitem{Bagnaschi:2011tu}
  E.~Bagnaschi, G.~Degrassi, P.~Slavich and A.~Vicini,
  %``Higgs production via gluon fusion in the POWHEG approach in the SM and in the MSSM,''
  arXiv:1111.2854 [hep-ph].
  %%CITATION = ARXIV:1111.2854;%%

\end{thebibliography}


\end{document}





